\section{Discussion}
\label{sec:discussion}

\subsection{Limitations}
\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=1pt]
  \item The stability study indicates that the other studies should filter for unstable
    answers: within both Qwen and Llama, the small model is $0\%$ format-invariant while
    the large model is $100\%$, so cosmetic format alone can flip a small model's answer.
  \item We do not have clean resolution between the small and large model on every axis;
    the sweep mixes families and sizes.
  \item The dynamics study is a case study of a single sample, not a population.
  \item Coverage is Spanish-only and four bias categories; the \textbf{genero} category
    is underpowered ($n{=}18$ total).
\end{itemize}

\paragraph{Dual-use and misuse risks.}
\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=1pt]
  \item Smaller models substitute confident stereotyped guessing for abstention, so a
    deployer choosing the cheapest model inherits the most biased behavior.
  \item Abstention can be gamed by answering ``unknown'' indiscriminately; the
    per-condition design scores indiscriminate abstention at $0$ on disambiguated items.
\end{itemize}

\subsection{Future Work}
\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=1pt]
  \item Study debiasing prompt scaffolds, which the title anticipates and we leave to
    future work.
  \item See how interventions look in activation space, not just at the output.
  \item Steering: turn the directions we find into causal, additive interventions.
\end{itemize}
