\section{Limitations and Dual-Use Considerations}
\label{sec:limitations}

\paragraph{Limitations.}
\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=1pt]
  \item \textbf{Single primary model.} The baseline, stability, selection,
    divergence, and geometry studies all run on one small model
    (\texttt{Qwen3-0.6B}); only the cross-model sweep spans five models (each
    family contributing two sizes), so the per-condition and geometric findings are
    not yet shown to generalize across models or scales.
  \item \textbf{Small per-cell $n$ on the pilot studies.} The stability and
    divergence studies use small item pools (stability $n{=}8/16$ items per
    condition; divergence $43$ scored items), so per-item Wilson CIs are wide and the
    slice-level rates (e.g.\ per-category abstention) should be read as pilot
    estimates rather than precise measurements. The selection / full-data run
    (\Cref{app:fulldata}) is the exception---$1347$ ambiguous items per scaffold.
  \item \textbf{Per-scaffold comparison on one model.} The full-data run
    (\Cref{app:fulldata}) scores all four \Cref{tab:scaffolds} scaffolds on $1347$
    ambiguous items each and selects $s^\star{=}$\texttt{interpretive\_direction}
    behaviorally ($99.9\%$ vs.\ $85.6\%$ baseline), corroborated by the
    representational ($+0.48$ silhouette, \Cref{app:geometry}) and causal
    (\Cref{app:steer}) findings; but this per-scaffold ranking is established only on
    \texttt{Qwen3-0.6B} and is not yet shown to hold across models or scales.
  \item \textbf{Spanish-only scope.} Coverage is Spanish (with matched English
    scaffold text) and four bias categories (racism, xenophobia, classism, gender);
    the \textbf{genero} category is underpowered ($n{=}18$ total; $n{=}6$ per
    disambiguated slice), and findings may not transfer to other languages or
    social-bias taxonomies.
  \item \textbf{Correlational geometry.} The geometry study measures which item
    labels separate the answer representation; it is correlational, not causal, and
    does not establish that the scaffold axis \emph{drives} abstention behavior.
  \item \textbf{Parser dependence.} The greedy- and sampled-thinking readouts depend
    on a parser that can drop free-form draws ($4$ of $47$ divergence prompts yielded
    no parsable thinking draw), which biases the parsed count $M$ and the thinking
    distribution $p^{\mathrm{th}}$.
  \item \textbf{Symmetry assumptions.} The position$\to$role remap and bilingual
    rendering assume the model treats display positions and the two languages
    symmetrically; residual position or language bias would leak into the reported
    rates.
\end{itemize}

\paragraph{Dual-use and misuse risks.}
\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=1pt]
  \item The debiasing scaffold is invertible: the one-sentence preamble that pushes
    the model toward ``unknown'' can be rewritten into a bias-amplifying preamble
    that pushes it toward the stereotyped target group.
  \item Small models substitute confident stereotyped guessing for abstention, so a
    deployer choosing the cheapest model inherits the most biased behavior.
  \item Abstention can be gamed by answering ``unknown'' indiscriminately; the
    per-condition design scores indiscriminate abstention at $0$ on disambiguated
    items.
\end{itemize}

\paragraph{Suggestions for future improvements.}
\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=1pt]
  \item Replicate the full-data per-scaffold ranking
    (\Cref{app:fulldata}: $s^\star{=}$\texttt{interpretive\_direction}, with
    \texttt{intent\_and\_register\_respect} below the no-scaffold baseline) across the
    cross-model sweep, to test whether the same scaffold wins at other scales.
  \item Extend the per-condition and geometry studies across the cross-model sweep's
    models and sizes to test whether the findings generalize beyond \texttt{Qwen3-0.6B}.
  \item Extend the causal steering of \Cref{app:steer}---which already turns the
    correlational scaffold axis into an additive intervention that raises abstention
    on held-out items---across more models, layers, and the disambiguated condition.
  \item Broaden coverage beyond Spanish and the four bias categories, and add items
    to the underpowered \textbf{genero} slice.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Code and Data}
\begin{itemize}[label=\textendash,leftmargin=*,topsep=2pt]
  \item \textbf{Code repository:} the evaluation harness (three-/two-option
    choice runners, greedy- and sampled-thinking readouts, scaffolds, and the
    geometry analysis) will be released with the final version.
  \item \textbf{Data/Datasets:} \textsc{SESGO}~\citep{robles2025sesgo}, used under
    its original license.
  \item \textbf{Models:} \texttt{Qwen3-0.6B} / \texttt{Qwen3-4B},
    \texttt{Llama-3.2-1B} / \texttt{-3B}, and \texttt{Mistral-7B-Instruct-v0.3},
    run locally.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% REFERENCES -- add entries to references.bib and cite with \citep{key}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{plainnat}
\bibliography{references}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% APPENDICES -- one per pipeline. Each appendix is, in order:
%%   (a) \paragraph{Methods.} -- pipeline-specific methodology
%%   (b) \paragraph{Results.} -- terse bulleted factual results
%%   (c) the figures for that pipeline (guarded \plotfig / \plotsub includes)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
