%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The Global South AI Safety Hackathon (June 2026) -- submission template
%% Single-column arXiv-style preprint. Build with: bash build.sh
%%
%% HOW TO WRITE IN THIS TEMPLATE
%%   - Guidance text is wrapped in \guide{...} and renders gray+italic.
%%     Replace each \guide{...} block with your own content, then remove it.
%%   - The yellow info box and all \guide{...} text MUST be deleted before
%%     submitting (see the box on page 1).
%%   - Recommended length: 4 pages excluding references and appendix.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[11pt]{article}

\input{paperdefs-common}
\begin{document}
\thispagestyle{plain}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% TITLE BLOCK
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent\rule{\textwidth}{1.2pt}\par
\vspace{0.8em}
\begin{center}
  {\LARGE\bfseries Scale Shapes Bias in Spanish-Prompted Open-Weight LLMs\par}
\end{center}
\vspace{0.6em}
\noindent\rule{\textwidth}{0.4pt}\par
\vspace{1.4em}

%% Author
\begin{center}
  {\LARGE Ian Rios-Sialer%
    \footnote{Research conducted at the \href{https://www.apartresearch.com/}{Global South AI Safety Hackathon} (Apart Research $\times$ Schmidt Sciences), June 19--21, 2026.}\par}\vspace{0.45em}
  \href{mailto:ian@unrulyabstractions.com}{ian@unrulyabstractions.com}\\[0.3em]
  Independent Researcher\\[0.3em]
  San Francisco, CA, USA\\[2.0em]
  With \href{https://www.apartresearch.com/}{Apart Research}
\end{center}
\vspace{1.2em}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% ABSTRACT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Large language models encode and amplify human social bias, and the harm falls
hardest on minoritized groups. Mitigating that harm needs interventions tuned to
each social context, yet until recently there was no Spanish-language data to even
measure such bias. We characterize bias in Spanish-prompted, open-weight LLMs across
model scale, using behavioral and distributional studies on the new \textsc{SESGO}
benchmark, where an item is \emph{ambiguous} when its text names no specific person (the
correct answer is \emph{unknown}, abstain) and \emph{disambiguated} when it specifies a
group. We report two preliminary findings. First, smaller models are more biased: across
a sweep of open-weight models, abstention accuracy rises with scale and the error split toward the
marginalized group narrows. Second, reasoning commits at a few discrete tokens: forking the
chain of thought of the thinking models on a bias-revealing item, every model from $0.8$B to
$9$B resolves its answer at a small number of decisive forking tokens rather than accreting it
gradually, a concrete locus for inspection or intervention. These results give future
LATAM-context interventions, including the debiasing prompt scaffolds we leave to future
work, a map of where bias lives, and we release the evaluation harness.
\end{abstract}

\bigskip

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{sections/introduction}
\input{sections/stability}
\input{sections/bias}
\input{sections/uncertainty}
\input{sections/conclusion}
\input{sections/code_and_data}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% REFERENCES -- add entries to references.bib and cite with \citep{key}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{plainnat}
\bibliography{references}

\section*{LLM Usage Statement}
We used an LLM assistant to help draft and format this report and to formalize
notation from the codebase. All design decisions, claims, and reported results
are verified by the authors against the harness implementation and the raw
\texttt{response\_samples.json} via the project's own slicing code.

\end{document}
